---
title: "dplyr::across()"
author: "shaggycamel"
date: December 5, 2020
output:
  html_document:
    df_print: paged
---

<br>
This article showcases an example of summarising a dataframe using a combination of
the `dplyr::summarise` and `dplyr::across` functions.

Within the `summarise` function, it is possible to summarise numerous columns that
adhere to any condition, using clean and succinct syntax with the help of the `across`
function.

Let's begin by loading the required packages...
```{r message=FALSE}
# "tidyverse" is a collection of packages that make up the tidy-universe
library(tidyverse)

# Packages included in tidyverse are:
tidyverse::tidyverse_packages(include_self = FALSE)
```
<br>
Most of the functions used in this article belong to the `dplyr` package.

The example dataset used in this article is the `ggplot2::diamonds` dataframe, it contains information of around 54,000 diamonds.

Numeric columns in this dataframe are:
<br>

```{r}
ggplot2::diamonds %>% select(where(is.numeric))
```
<br>
For each numeric column, the *min*, *max* and *mean* can be calculated for each group in the **cut** column with help of the across function. Summarisation is as follows:
<br>
```{r message=FALSE}
# With the diamonds dataframe:
(df <- ggplot2::diamonds %>% 
  
  # for each category in the "cut" column
  dplyr::group_by(cut) %>% 
  
  # calculate summaries for...
  dplyr::summarise(
    
    # ...each column that:
    dplyr::across(
      
      # is numeric
      .cols = where(is.numeric)
      
      # returning the max, min and mean value
      , .fns = c(min, max, mean)
      
      # and name columns accordingly
      , .names = "{.fn}_{.col}"
    )
  )
)
```
<br>
As can be seen in the above output, column names do not reflect the summarisation function used. Instead, column names follow the logic **1_\<attribute\>**, **2_\<attribute\>**, **3_\<attribute\>** where **1**, **2** and **3** represent 
the *min*, *max* and *mean*, respectively.

Here is a workaround to fix column names:
<br>
```{r}
colnames(df) <- colnames(df) %>% 
  stringr::str_replace_all(
    pattern = c(
      "1" = "min"
      , "2" = "max"
      , "3" = "mean"
    )
  )

df
```
<br>
Voilà! The *min*, *max* and *mean* for each **cut** group have been calculated for each numeric column, complete with sensible column names, within the diamonds dataframe.

<br>
As shown in this article, the across function allows one to carryout summarisation on many columns at once. This is very handy when wanting to summarise a large dataframe. 

It is also possible for columns in the `across` function to be referenced by a naming convention. For example, if one wishes to summarise all columns in a dataset that end in a particular word, it can be achieved using the following:
<br>
```{r eval=FALSE}
dplyr::across(
  
  # ends with "word"
  .cols = dplyr::ends_with("word")
  
  # returning the mean
  , .fns = mean
)
```
<br>
In addition to `across` complimenting `summarise` so well, it can also be used in conjunction with `dplyr::mutate` and `dplyr::select`.

Why not try out the across function today !?

:D
<br>
<br>
<br>
<br>


